---
title: "Analysis"
author: "Matthew Loder"
date: "7/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning scripts:
```{r packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(gridExtra)
library(DescTools)
library(splines)
library(stringr)
source("helpers.R")


theme_set(theme_bw())

colscale = scale_color_manual(values=c("#7CB637", "#666666", "#4381C1")) 
colscale_fill = scale_fill_manual(values=c("#7CB637", "#666666", "#4381C1")) 

remove_quotes = function(d) {
  
}

format_data = function(d) {
  drops <- c("modal1","rating1")
  d2 = d[ , !(names(d) %in% drops)]
  setnames(d2, old=c("rating2","modal2"), new=c("rating", "modal"))
  
  drops <- c("modal2","rating2")
  d3 = d[ , !(names(d) %in% drops)]
  setnames(d3, old=c("rating1","modal1"), new=c("rating", "modal"))
  
  drops <- c("modal2", "rating2", "modal1", "rating1")
  d4 = d[ , !(names(d) %in% drops)]
  d4$rating = d4$rating_other
  d4$modal = "other"
  
  d = rbind(d2, d3, d4)
  
  d$modal = gsub('"', '', d$modal)
  d$modal = factor(d$modal, levels = c("might", "probably", "other"), ordered = T)
  
  d$percent_window_f = factor(d$percent_window)
  
  d_aisle = d %>% filter(.,grepl("aisle", type))
  d_window = d %>% filter(.,grepl("window", type))
  
  d_aisle_reverse = d_aisle
  d_aisle_reverse$percent_window = 100-d_aisle$percent_window
  
  d_comparison = rbind(d_window, d_aisle_reverse)
  d_comparison$percent_window_f = factor(d_comparison$percent_window)
  
  return(d_comparison)
}

format_catch_trials = function(d) {
  
  if (d$condtion == '1' || d$condtion == '2') {
    d_window_60 = d %>% filter(.,grepl("60", image))
    d_window_25 = d %>% filter(.,grepl("25", image))
    d_window_100 = d %>% filter(.,grepl("100", image))
    d_aisle_25 = d %>% filter(.,grepl("75", image))
    d_aisle_60 = d %>% filter(.,grepl("40", image))
    
    d_window_60$prob = 60
    d_aisle_60$prob = 60
    d_window_100$prob = 100
    d_window_25$prob = 25
    d_aisle_25$prob = 25
    
    d_output = rbind(d_window_60, d_window_25, d_window_100, d_aisle_25, d_aisle_60)
  } else if (d$condtion == '3' || d$condtion == '4') {
    d_window_60 = d %>% filter(.,grepl("60", image))
    d_window_90 = d %>% filter(.,grepl("90", image))
    d_window_100 = d %>% filter(.,grepl("100", image))
    d_aisle_90 = d %>% filter(.,grepl("10", image))
    d_aisle_60 = d %>% filter(.,grepl("40", image))
    
    d_window_60$prob = 60
    d_window_90$prob = 90
    d_window_100$prob = 100
    d_aisle_60$prob = 60
    d_aisle_90$prob = 90
    
    d_output = rbind(d_window_60, d_window_90, d_window_100, d_aisle_90, d_aisle_60)
  } 
  
  return(d_output)
}

```

## Load data:

```{r data, echo=FALSE}
trials_cond1 = format_data(read.csv("../data/04_exp_away_cond1-trials.csv"))
trials_cond2 = format_data(read.csv("../data/04_exp_away_cond2-trials.csv"))
trials_cond3 = format_data(read.csv("../data/04_exp_away_cond3-trials.csv"))
trials_cond4 = format_data(read.csv("../data/04_exp_away_cond4-trials.csv"))

catch_trials1 = format_catch_trials(read.csv("../data/04_exp_away_cond1-exp_trials.csv"))
catch_trials2 = format_catch_trials(read.csv("../data/04_exp_away_cond2-exp_trials.csv"))
catch_trials3 = format_catch_trials(read.csv("../data/04_exp_away_cond3-exp_trials.csv"))
catch_trials4 = format_catch_trials(read.csv("../data/04_exp_away_cond4-exp_trials.csv"))

d1 = trials_cond1
d2 = trials_cond2
d3 = trials_cond3
d4 = trials_cond4

# combine data: 
d1$Answer.condition = "optimist"
d2$Answer.condition = "confident"
d3$Answer.condition = "pessimist"
d4$Answer.condition = "cautious"
```


## Exclude random responses

```{r}
exclude_random = function(d) {
  d_overall_means = d %>%
  group_by(modal, workerid) %>% 
  summarise(rating_m_overall = mean(rating))

  d_indiv_means =  d %>%
    group_by(modal,percent_window, workerid) %>% 
    summarise(rating_m = mean(rating))
  
  d_indiv_merged = merge(d_indiv_means, d_overall_means, by=c("workerid", "modal"))
  
  cors = d_indiv_merged %>%
    group_by(workerid) %>%
    summarise(corr = cor(rating_m, rating_m_overall))
  
  exclude = cors %>%
    filter(corr > 0.75) %>%
    .$workerid
  
  print(paste("Excluded", length(exclude), "participants based on random responses."))
  
  d = d %>% filter(!(workerid %in% exclude))
}

d1 = exclude_random(d1)
d2 = exclude_random(d2)
d3 = exclude_random(d3)
d4 = exclude_random(d4)
```

## Aggregated results

```{r aggregate_plots, echo=FALSE, fig.width=10, fig.height=8}
plot_conditions = function(d, plot_title) {
  d_means = d %>%
    group_by(workerid, percent_window, modal) %>%
    summarise(participant_mean = mean(rating)) %>%
    group_by(percent_window, modal) %>%
    summarise(mu = mean(participant_mean),
              ci_high = ci.high(participant_mean),
              ci_low = ci.low(participant_mean))

  p1 = ggplot(d_means, aes(x = percent_window, y = mu, col=modal)) + 
    xlab("% preferred seating") +
    ylab("mean ratings") +
    geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
    geom_line() +
    geom_point(size=1) +
    guides(col=guide_legend(title="Expr.")) +
    colscale +
    theme(legend.position="bottom") + 
    ggtitle(plot_title) 
  
  p2 = ggplot(d, aes(x=percent_window_f, y=rating, fill=modal)) + 
    geom_boxplot() +
    ggtitle(plot_title) + xlab("percentage") + colscale_fill +   theme(legend.position="bottom")
  
  by_participant = ggplot( d %>%
    group_by(workerid, percent_window, modal) %>% 
    summarise(participant_mean = mean(rating)) , 
      aes(x=percent_window, y=participant_mean, col=modal)
        ) + 
    xlab("% preferred seat") +
    ylab("mean ratings") +
    geom_line() +
    geom_point(size=1) +
    guides(col=guide_legend(title="Expr.")) +
    colscale +
    theme(legend.position="bottom") + ggtitle(plot_title) + facet_wrap(~workerid, ncol = 8)
    return(list("p1" = p1, "p2" = p2, "by_participant" = by_participant))
}

plot_2conditions = function(d1, d2, plot_title) {
  
  d2$workerid = d2$workerid + max(d1$workerid) + 1
  
  d = rbind(d1,d2)
  d$condition = factor(d$condition)
  
  d_means = d %>%
    group_by(workerid, percent_window, modal, condition) %>%
    summarise(participant_mean = mean(rating)) %>%
    group_by(percent_window, modal, condition) %>%
    summarise(mu = mean(participant_mean),
              ci_high = ci.high(participant_mean),
              ci_low = ci.low(participant_mean))

  p1 = ggplot(d_means, aes(x = percent_window, y = mu, col=modal, lty=condition)) + 
    xlab("% preferred seating") +
    ylab("mean ratings") +
    geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
    geom_line() +
    geom_point(size=1) +
    guides(col=guide_legend(title="Expr.")) +
    colscale +
    theme(legend.position="bottom") + 
    ggtitle(plot_title) 
  
  p2 = ggplot(d, aes(x=percent_window_f, y=rating, fill=modal)) + 
    geom_boxplot() +
    ggtitle(plot_title) + xlab("percentage") + colscale_fill +   theme(legend.position="bottom")
  
  by_participant = ggplot( d %>%
    group_by(workerid, percent_window, modal) %>% 
    summarise(participant_mean = mean(rating)) , 
      aes(x=percent_window, y=participant_mean, col=modal)
        ) + 
    xlab("% preferred seat") +
    ylab("mean ratings") +
    geom_line() +
    geom_point(size=1) +
    guides(col=guide_legend(title="Expr.")) +
    colscale +
    theme(legend.position="bottom") + ggtitle(plot_title) + facet_wrap(~workerid, ncol = 8)
    return(list("p1" = p1, "p2" = p2, "by_participant" = by_participant))
}

ps1 = plot_conditions(d1, "Optimist")
ps2 = plot_conditions(d2, "Confident")
ps3 = plot_conditions(d3, "Pessimist")
ps4 = plot_conditions(d4, "Cautious")

plot(ps1$p1)
plot(ps2$p1)
plot(ps3$p1)
plot(ps4$p1)

ps24 = plot_2conditions(d2,d4, "Confident-cautious")
plot(ps24$p1)

ps13 = plot_2conditions(d1,d3, "Optimist-pessimist")
ps12 = plot_2conditions(d1,d2, "Optimist-confident")
ps34 = plot_2conditions(d3,d4, "Pessimist-cautious")
plot(ps13$p1)
plot(ps12$p1)
plot(ps34$p1)

```
## Individual plots

```{r by_participant_plots}
plot(ps1$by_participant)
plot(ps2$by_participant)
plot(ps3$by_participant)
plot(ps4$by_participant)
```

## AUC Computation

```{r auc_method_1, echo=FALSE}

auc_method1 = function(d) {
      #estimate linear model
    model = lm(formula = rating_m ~ ns(percent_window, df = 4), data = d)
  
    pred = data.frame(0:1000)
    pred$percent_window = (0:1000)/10
    pred$rating = predict(model, pred)
    auc = AUC(x = pred$percent_window, y=pred$rating)
    return(auc)
}

auc_for_participants = function(d, method) {
  
  aucs = d %>% 
    group_by(workerid) %>%
    summarize(auc_might = 0, auc_probably = 0)
  
i = 1

  for (wid in unique(d$workerid)) {
    d.might_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "might") %>%
      group_by(workerid, percent_window) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_might[i] = method(d.might_ratings)
    
    d.probably_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "probably") %>%
      group_by(workerid, percent_window) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_probably[i] = method(d.probably_ratings)
    i = i + 1
  }
  
  aucs$auc_diff = aucs$auc_might - aucs$auc_probably
  
  return(aucs)
}

aucs.optimist = d1 %>% auc_for_participants(., method=auc_method1)

aucs.confident = d2 %>% auc_for_participants(., method=auc_method1)

aucs.pessimist = d3 %>% auc_for_participants(., method=auc_method1)

aucs.cautious = d4 %>% auc_for_participants(., method=auc_method1)

aucs.optimist$cond = "Optimistic condition"

aucs.confident$cond = "Confident condition"

aucs.pessimist$cond = "Pessimistic condition"

aucs.cautious$cond = "Cautious condition"

aucs.all1 = rbind(aucs.optimist, aucs.confident)

aucs.all2 = rbind(aucs.pessimist, aucs.cautious)

aucs.all3 = rbind(aucs.optimist, aucs.confident, aucs.pessimist, aucs.cautious)

auc.all4 = rbind(aucs.optimist, aucs.pessimist)

aucs.all1 = aucs.all1 %>% 
  group_by(., cond) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))

ggplot(aucs.all1, aes(x=0, y=auc_diff_m, color=cond)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 

aucs.all2 = aucs.all2 %>% 
  group_by(., cond) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))

ggplot(aucs.all2, aes(x=0, y=auc_diff_m, color=cond)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 

aucs.all3 = aucs.all3 %>% 
  group_by(., cond) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))

ggplot(aucs.all3, aes(x=0, y=auc_diff_m, color=cond)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 
```

So, we're seeing what we expected for the confident speaker, with the AUC > in the optimistic condition than in the confident condition (indicating explaining away), but we're not seeing the same (with just one speaker) for the pessimistic/cautious condition, where we would expect flipped results, with the pessimistic condition having a lower AUC than the cautious condition (with adaptation). To sum up, we expect the order to be, from greater to lower difference, "cautious", "pessimistic", "optimistic", and "confident".
```{r t_test_method1, echo=FALSE}

print(t.test(aucs.confident$auc_diff, aucs.cautious$auc_diff, var.equal=TRUE))

print(t.test(aucs.pessimist$auc_diff, aucs.cautious$auc_diff, var.equal=TRUE))



```

## Power analysis
```{r}
m.cautious = mean(aucs.cautious$auc_diff)
m.pessimist = mean(aucs.pessimist$auc_diff)

sd.cautious = sd(aucs.cautious$auc_diff)
sd.pessimist = sd(aucs.pessimist$auc_diff)

d = m.cautious - m.pessimist
d = d/sd.pessimist

d

power.t.test(n=60, delta=d, sd=1, sig.level = 0.05, type="two.sample", alternative="two.sided")

```

