---
title: "Analysis of speaker-specific adaptation experiment"
author: "Matthew Loder and Sebastian Schuster"
date: "07/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(gridExtra)
library(DescTools)
library(splines)
library(stringr)
source("helpers.R")

theme_set(theme_bw())


colscale = scale_color_manual(values=c("#7CB637", "#666666", "#4381C1")) 
colscale_fill = scale_fill_manual(values=c("#7CB637", "#666666", "#4381C1")) 

remove_quotes = function(d) {
  if (!is.null(d$modal)) {
    d$modal = gsub('"', '', d$modal)
  }
  d$color = gsub('"', '', d$color)
  d$speaker_type = gsub('"', '', d$speaker_type)
  return(d)
}

format_data = function(d) {
  
    drops <- c("modal1","rating1")
  d2 = d[ , !(names(d) %in% drops)]
  setnames(d2, old=c("rating2","modal2"), new=c("rating", "modal"))
  
  drops <- c("modal2","rating2")
  d3 = d[ , !(names(d) %in% drops)]
  setnames(d3, old=c("rating1","modal1"), new=c("rating", "modal"))
  
  drops <- c("modal2", "rating2", "modal1", "rating1")
  d4 = d[ , !(names(d) %in% drops)]
  d4$rating = d4$rating_other
  d4$modal = "other"
  
  d = rbind(d2, d3, d4)
  
  d$modal = gsub('"', '', d$modal)
  d$modal = factor(d$modal, levels = c("might", "probably", "other"), ordered = T)
  
  d$percentage_blue_f = factor(d$percentage_blue)
  d_middle = d %>% filter(., grepl("blue", sentence2))
  d_aislewindow = d %>% filter(., grepl("aisle", sentence2))
  
  d_orange_reverse = d_orange
  d_orange_reverse$percentage_blue = 100-d_orange$percentage_blue
  
  d_comparison = rbind(d_blue, d_orange_reverse)
  d_comparison$blue= grepl("blue", d_comparison$sentence2)
  d_comparison$percentage_blue_f = factor(d_comparison$percentage_blue)
  return(d_comparison)
  
}
```

```{r data, echo=FALSE}
# Load data
trials = remove_quotes(format_data(read.csv("../data/11_talker_specific_adaptation_fixed-trials.csv")))
exp_trials = remove_quotes(read.csv("../data/11_talker_specific_adaptation_fixed-exp_trials.csv"))
trials[, c("test_order", "first_speaker_type", "confident_speaker")] = str_split(trials$condition, "_", simplify=T)
```

## Catch trial performance

```{r catch_trials, echo=FALSE}
get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  filter(., catch_trial == 1) %>%
  group_by(workerid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct))
  
  return(ret)
}
EXCLUDE_BELOW = 11
catch_trial_perf = get_correct_catch_trial_counts(exp_trials)
exclude = catch_trial_perf %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid
print(paste("Excluded", length(exclude), "participants based on catch-trial performance."))
#final data
d = trials %>% filter(., !(workerid %in% exclude))
```

```{r perf_plot_1, fig.height=2,fig.width=10, echo=FALSE}
cutoff <- data.frame(yintercept=EXCLUDE_BELOW-0.5, cutoff=factor(EXCLUDE_BELOW-0.5))
ggplot(catch_trial_perf, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")
```

## Exclusion of random guesses

We further exclude participants who seem to provide random ratings independent of the scene that they are seeing. We quantify this by computing the mean rating for each utterance across all trials for each participant and computing the correlation between a participant's actual ratings and their mean rating. A high correlation is unexpected and indicates that a participant chose ratings at random. We therefore also exclude the data from participants for whom this correlation is larger than 0.75.

```{r random_guess_exclusions, echo=FALSE}
d_overall_means = d %>%
  group_by(modal, workerid) %>% 
  summarise(rating_m_overall = mean(rating))
d_indiv_means =  d %>%
  group_by(modal,percentage_blue, workerid) %>% 
  summarise(rating_m = mean(rating))
d_indiv_merged = merge(d_indiv_means, d_overall_means, by=c("workerid", "modal"))
cors = d_indiv_merged %>%
  group_by(workerid) %>%
  summarise(corr = cor(rating_m, rating_m_overall))
exclude = cors %>%
  filter(corr > 0.75) %>%
  .$workerid
print(paste("Excluded", length(exclude), "participants based on random responses."))
d = d %>% filter(!(workerid %in% exclude))
```


## Aggregated results


```{r aggregate_plots, echo=FALSE, fig.width=10, fig.height=8}
plot_conditions = function(d, plot_title) {
  
  
  d_means = d %>%
  group_by(workerid, percentage_blue, modal, speaker_type) %>% 
  summarise(participant_mean = mean(rating)) %>%
  group_by(percentage_blue, modal, speaker_type) %>%
  summarise(mu = mean(participant_mean),
         ci_high = ci.high(participant_mean), 
         ci_low = ci.low(participant_mean))
  
  p1 = ggplot(d_means, aes(x=percentage_blue, y=mu, col=modal)) + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expr.")) +
  colscale +
  theme(legend.position="bottom") + 
  ggtitle(plot_title) +
  facet_wrap(~speaker_type)
  p2 = ggplot(d, aes(x=percentage_blue_f, y=rating, fill=modal)) + 
  geom_boxplot() +
  ggtitle(plot_title) + xlab("percentage") + colscale_fill +   theme(legend.position="bottom")
  by_participant = ggplot( d %>%
  group_by(workerid, percentage_blue, modal, speaker_type) %>% 
  summarise(participant_mean = mean(rating)) , 
    aes(x=percentage_blue, y=participant_mean, col=modal, lty=speaker_type)
  ) + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expr.")) +
  colscale +
  theme(legend.position="bottom") + ggtitle(plot_title) + facet_wrap(~workerid, ncol = 8)
  return(list("p1" = p1, "p2" = p2, "by_participant" = by_participant))
}
  ps1 = plot_conditions(d, "")
  plot(ps1$p1)
```

## Individual responses

```{r indiv_plots, echo=FALSE, fig.width=10, fig.height=20}
    
  plot(ps1$by_participant)
```

## AUC computation

We use the `AUC` function with the `splines` method to directly compute the AUC.

```{r auc_method_1, echo=FALSE}
auc_method2 = function(d) {
  auc = AUC(x=d$percentage_blue, y=d$rating_m, method="spline")
  return(auc)
}
auc_for_participants = function(d, method) {
  
  d[d$color=="orange",]$percentage_blue = 100 - d[d$color=="orange",]$percentage_blue 
  
  aucs = d %>% 
    group_by(workerid) %>% 
    summarize(test_order = first(test_order),
              first_speaker_type = first(first_speaker_type),
              confident_speaker = first(confident_speaker))
    
    
  aucs$auc_might = 0
  aucs$auc_probably = 0
  
  i = 1
  
  for (wid in unique(d$workerid)) {
    d.might_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "might") %>%
      group_by(workerid, percentage_blue) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_might[i] = method(d.might_ratings)
    d.probably_ratings = d %>% 
      filter (workerid == wid) %>%
      filter (modal == "probably") %>%
      group_by(workerid, percentage_blue) %>%
      summarise(rating_m = mean(rating))
    
    aucs$auc_probably[i] = method(d.probably_ratings)
    i = i + 1
  }
  
  aucs$auc_diff = aucs$auc_might - aucs$auc_probably
  
  return(aucs)
}
#AUCs for cautious speaker ratings 
aucs.cautious = d %>% filter(speaker_type == "cautious") %>% auc_for_participants(., method=auc_method2)
#AUCs for confident speaker ratings 
aucs.confident = d %>% filter(speaker_type == "confident") %>% auc_for_participants(., method=auc_method2)
```


```{r auc_method1_plots, fig.width=4, fig.height=4, echo=FALSE}
aucs.cautious$cond = "cautious (might-biased)"
aucs.confident$cond = "confident (probably-biased)"
aucs.all = rbind(aucs.cautious, aucs.confident)
aucs.all = aucs.all %>% 
  group_by(., cond) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))
ggplot(aucs.all, aes(x=0, y=auc_diff_m, color=cond)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 
```

t-test and regression model with control variables:

```{r t_test_method1, echo=FALSE}
print(t.test(aucs.cautious$auc_diff, aucs.confident$auc_diff, var.equal=TRUE))
model = lm(formula= auc_diff ~ cond + test_order + first_speaker_type + confident_speaker, data = rbind(aucs.cautious, aucs.confident))
print(summary(model))
```

